IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2020-10-10 main.tex
:ambulance: Fix convergence proof

		\begin{subsection}{IRS-aided SWIPT}
			The effective channel enhancement and low power consumption of IRS are expected to bring more opportunities to SWIPT. It was argued in \cite{Wu2019b} that dedicated energy beam is not required to maximize the Weighted Sum-Power (WSP) for a multiuser IRS-assisted SWIPT system, which is power-inefficient and may cause interference. On the other hand, \cite{Tang2019} suggested that if the interference from energy signals can be cancelled, multiple energy beams are generally required for the max-min harvested power problem. \cite{Wu2019c} proposed a novel penalty-based algorithm, whose inner layer employs Block Coordinate Descent (BCD) method to update transmit precoders and IRS phase shifts while the outer layer updates the penalty coefficients. It demonstrated that Line-of-Sight (LoS) links can boost the harvested power because the rank-deficient channels are highly correlated. In such cases, a single energy stream can satisfy the energy constraints of all energy receivers.

			However, most existing IRS-assisted SWIPT papers relied on an oversimplified linear harvester model with focus on single-carrier transmission for separated information and energy receivers. Rectifier nonlinearity and co-localized receiver, acknowledged as key to practical and efficient SWIPT system design, have not yet been addressed in the relevant research. In this paper, we depart from a tractable harvester model and study an IRS-aided downlink MISO SWIPT system, where the user employs the received signal for both information decoding and energy harvesting. The contributions are as follows.

			In this paper, we marry the benefits of joint waveform and active beamforming optimization for SWIPT (accounting for nonlinearity) with the passive beamforming capability of IRS. The contribution of the paper are listed as follows. First, we introduce an novel IRS-aided SWIPT architecture based on a joint waveform, active and passive beamforming design. This architecture can be seen to IRS-aided SWIPT what [40] is to SWIPT. Specifically, we consider a multi-carrier IRS-aided downlink MISO SWIPT system where the IRS assists the information and energy transmission to a single user. A multicarrier unmodulated power waveform (deterministic multisine) is superposed to a multicarrier modulated information waveform to boost the energy transfer efficiency. The power and information waveforms and active beamforming at the transmitter, the phase shifts at the IRS and the splitting ratio at the receiver are jointly optimized to maximize the R-E tradeoff while accounting for the rectenna nonlinearity. This is the first paper to proposed such a joint waveform, active and passive beamforming architecture for IRS-aided SWIPT. Note that existing IRS-aided SWIPT papers [36,37,38,39] focus on single-carrier transmissions and ignore the rectenna nonlinearity (and therefore assume the oversimplified and inaccurate linear model), which prevents from exploiting the waveform gain in SWIPT system design. On the contrary, this paper focus on multicarrier IRS-SWIPT and investigates the fundamental impact of harvester nonlinearity on joint waveform, active and passive beamforming design. Second, we formulate an optimization for this joint waveform, active and passive beamforming design so as to characterize the R-E region achieved by the proposed architecture. The R-E region characterization problem is transformed into multiple energy maximization problems subject to different rate constraints. To reduce the design complexity, we propose an Alternating Optimization (AO) algorithm that updates the channel and transceiver [Not clear yet what you mean by this, maybe you can be more explicit and say that you optimize alternatively the IRS weights (passive beamforming) and the waveform and active beamforming weights] iteratively based on Semidefinite Relaxation (SDR) and Geometric Programming (GP) techniques. [Here I would add one sentence that says: "In contrast to [...], ... " to explain how th optimization of IRS-aided SWIPT changes compared to SWIPT]. Third, we conduct numerical evaluations that demonstrate the benefit of the proposed architecture, and that the proposed algorithm can find a stationary point for all tested channel realizations. We demonstrate that dedicated power waveform can boost the energy transmission efficiency such that TS and PS are preferred at low SNR and high SNR, respectively. Also, IRS is shown to bring a significant channel amplification and R-E enhancement especially when deployed next to the transmitter or the receiver. Finally, the proposed adaptive IRS-aided SWIPT architecture and design outperform the benchmark schemes for broadband transmission, and the optimal IRS can be approximated in closed form for single transmit antenna narrowband SWIPT [This last sentence is not clear to me].

@@

			\textit{First}, we extend the SWIPT waveform design in \cite{Clerckx2018b} to the IRS case and investigate the passive beamforming issue.

			% single antenna, DoF -> new words. ff irs fs channel not all flexibility; active antenna -> frequency

			We investigate the impact of IRS development on precoder design, waveform preference and transceiving strategy. Comparison is made with the no IRS case to highlight the benefit of introducing IRS in SWIPT systems.

			\textit{Second},

			Each achievable R-E pair is obtained through an Alternating Optimization (AO) algorithm that updates the IRS reflection coefficients

			and transceiver iteratively based on Semidefinite Relaxation (SDR) and Geometric Programming (GP) techniques.
		\end{subsection}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and different rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\ref{op:original} is intricate due to the non-convex objective function \ref{ob:original} and the rate constraint \ref{co:original_rate} with coupled variables. To reduce the design complexity, we propose a suboptimal AO algorithm that iteratively updates the IRS phase shift and transmit waveform plus the receive splitting ratio until convergence.

@@

		\begin{subsection}{Convergence}
			\begin{proposition}\label{pr:irs}
				For any feasible initial point, the proposed SCA-based Algorithm~\ref{al:irs} is guaranteed to converge to a stationary point of the IRS phase shift subproblem.
			\end{proposition}

			\begin{proof}\label{pf:irs}
				The objective function \ref{ob:irs} is non-decreasing over iterations because the solution of problem~\ref{op:irs} at iteration $i-1$ is still a feasible point at iteration $i$. Moreover, the sequence $\{\tilde{z}(\boldsymbol{\Phi}^{(i)})\}_{i=1}^{\infty}$ is bounded above due to the unit-modulus constraint \ref{co:irs_modulus}. Thus, Algorithm~\ref{al:irs} is guaranteed to converge. To prove $\boldsymbol{\Phi}^{(i)}$ converge to the set of stationary points of IRS subproblem, we notice that the SCA-based Algorithm~\ref{al:irs} is indeed an inner approximation algorithm \cite{Marks1978}, since $\tilde{z}(\boldsymbol{\Phi}) \le z(\boldsymbol{\Phi})$, $\partial\tilde{z}(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}=\partial z(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}$ and the approximation \ref{eq:taylor_1} -- \ref{eq:taylor_3} are asymptotically tight as $i \to \infty$ \cite{Adali2010,Li2013}. Therefore, Algorithm~\ref{al:irs} converges to a stationary point.
			\end{proof}

>>>>

		\begin{subsection}{IRS-aided SWIPT}
			The effective channel enhancement and low power consumption of IRS are expected to bring more opportunities to SWIPT. It was argued in \cite{Wu2019b} that dedicated energy beam is power-inefficient and may cause interference, thus not required to maximize the Weighted Sum-Power (WSP) for a multiuser IRS-assisted SWIPT system. On the other hand, \cite{Tang2019} suggested that if the interference can be cancelled, multiple energy beams are generally required for the max-min harvested power problem. \cite{Wu2019c} proposed a novel penalty-based algorithm, whose inner layer employs Block Coordinate Descent (BCD) method to update transmit precoders and IRS phase shifts while the outer layer updates the penalty coefficients. It demonstrated that Line-of-Sight (LoS) links can boost the harvested power because the rank-deficient channels are highly correlated. In such cases, a single energy stream is enough for all energy receivers.

			However, to the best of the authors' knowledge, all existing IRS-assisted SWIPT papers focus on single-carrier transmission based on an oversimplified linear harvester model for separated information and energy receivers. Rectifier nonlinearity and co-localized receiver, acknowledged as key to practical and efficient SWIPT system design, have not yet been addressed in the relevant research. In this paper, we marry the benefits of joint waveform and active beamforming optimization for SWIPT (accounting for nonlinearity) with the passive beamforming capability of IRS. The contribution of the paper are listed as follows.

			\textit{First}, we introduce an novel IRS-aided SWIPT architecture based on a joint waveform, active and passive beamforming design. This architecture can be seen to IRS-aided SWIPT what \cite{Clerckx2018b} is to SWIPT. Specifically, we consider a multi-carrier IRS-aided downlink MISO SWIPT system where the IRS assists the information and energy transmission to a single user. A multicarrier unmodulated power waveform (deterministic multisine) is superposed to a multicarrier modulated information waveform to boost the energy transfer efficiency. The power and information waveforms and active beamforming at the transmitter, the phase shifts at the IRS and the splitting ratio at the receiver are jointly optimized to maximize the R-E tradeoff while accounting for the rectenna nonlinearity. This is the first paper to proposed such a joint waveform, active and passive beamforming architecture for IRS-aided SWIPT. Note that existing IRS-aided SWIPT papers \cite{Wu2019b,Tang2019,Wu2019c,Pan2019a} focus on single-carrier transmissions and ignore the rectenna nonlinearity (and therefore assume the oversimplified and inaccurate linear model), which prevents from exploiting the waveform gain in SWIPT system design. On the contrary, this paper focus on multicarrier IRS-SWIPT and investigates the fundamental impact of harvester nonlinearity on joint waveform, active and passive beamforming design.

			\textit{Second}, we formulate an optimization for this joint waveform, active and passive beamforming design so as to characterize the R-E region achieved by the proposed architecture. The R-E region characterization problem is transformed into multiple energy maximization problems subject to different rate constraints. Each achievable R-E pair is obtained through an Alternating Optimization (AO) algorithm that iteratively updates 1) phase shift at the IRS 2) waveform and active beamforming at the transmitter, power splitting ratio at the receiver, until convergence. On top of the SWIPT optimization based on Geometric Programming (GP) \cite{Clerckx2018b}, our IRS-aided SWIPT optimization employs an low-complexity passive beamforming algorithm based on Successive Convex Approximation (SCA) and Semidefinite Relaxation (SDR), which solves the highly non-convex problem in a efficient and reliable manner. Despite no rigorous mathematical proof exists for the convergence of the relaxed IRS phase shift problem, numerical results demonstrated SDR is tight and the proposed algorithm can always find a stationary point for all tested channel realizations under different configurations.

			\textit{Third}, we conduct numerical evaluations that demonstrate the benefit of the proposed architecture. It is demonstrated that the observations for SWIPT carries on to IRS-aided SWIPT, namely 1)

			dedicated power waveform can boost the energy transmission efficiency such that TS and PS are preferred at low SNR and high SNR, respectively. Also, IRS is shown to bring a significant channel amplification and R-E enhancement especially when deployed next to the transmitter or the receiver. Finally, the proposed adaptive IRS-aided SWIPT architecture and design outperform the benchmark schemes for broadband transmission, and the optimal IRS can be approximated in closed form for single transmit antenna narrowband SWIPT [This last sentence is not clear to me].

@@

			% single antenna, DoF -> new words. ff irs fs channel not all flexibility; active antenna -> frequency
		\end{subsection}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and different rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\ref{op:original} is intricate due to the non-convex objective function \ref{ob:original} and the rate constraint \ref{co:original_rate} with coupled variables. To reduce the design complexity, we propose a suboptimal AO algorithm that iteratively updates the phase shift at the IRS, the waveform and active beamforming at the transmitter, and the power splitting ratio at the receiver, until convergence.

@@

		\begin{subsection}{Convergence}
			\begin{proposition}\label{pr:irs}
				For any feasible initial point, the proposed SCA-based Algorithm~\ref{al:irs} can provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfy the KKT conditions, although there is no guarantee $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}. (Numerical results proved that $\boldsymbol{\Phi}^{\star}$ is rank-\num{1} for all tested channel realizations under different configurations.)
			\end{proposition}

			\begin{proof}\label{pf:irs}
				The objective function \ref{ob:irs} is non-decreasing over iterations because the solution of problem~\ref{op:irs} at iteration $i-1$ is still a feasible point at iteration $i$. Moreover, the sequence $\{\tilde{z}(\boldsymbol{\Phi}^{(i)})\}_{i=1}^{\infty}$ is bounded above due to the unit-modulus constraint \ref{co:irs_modulus}. Thus, Algorithm~\ref{al:irs} is guaranteed to converge. To prove $\boldsymbol{\Phi}^{(i)}$ converge to the set of stationary points of IRS subproblem, we notice that the SCA-based Algorithm~\ref{al:irs} is indeed an inner approximation algorithm \cite{Marks1978}, since $\tilde{z}(\boldsymbol{\Phi}) \le z(\boldsymbol{\Phi})$, $\partial\tilde{z}(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}=\partial z(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}$ and the approximation \ref{eq:taylor_1} -- \ref{eq:taylor_3} are asymptotically tight as $i \to \infty$ \cite{Adali2010,Li2013}. Therefore, Algorithm~\ref{al:irs} is guaranteed to provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfy the KKT conditions.
			\end{proof}
